---
title: "Technical Guideline Series"
output:
  word_document:
    toc: yes
    toc_depth: '4'
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri("C:/Users/jkumagai/Documents/IPBES/R/Geoinformatics/Technical Guidelines Series/Contributing_authors_template/logo2.png"), 
               alt = 'logo', 
                style = 'position:absolute; top:0; right:0; padding:10px; width: 400px; height: 100px;') 
```
```{r, echo = FALSE}
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```

## Part 10 - Contributing authors template email considering ILK and the data and knowledge management policy

**Prepared by Joy Kumagai and Peter Bates**   
**Reviewed by Aidin Niamir, the Task Force on Knowledge and Data and the Task Force on Indigenous and Local Knowledge**

*For any inquires please contact [tsu.data\@ipbes.net](mailto:tsu.data@ipbes.net) and [ilk.tsu.ipbes\@unesco.org](mailto:ilk.tsu.ipbes@unesco.org)*

Version: 1.0   
Last Updated: 18 July 2022

DOI: [10.5281/zenodo.6655926](https://doi.org/10.5281/zenodo.6655926)

This technical guideline is a template invitation email for contributing authors to IPBES assessments, which covers aspects relating to ILK and the data and knowledge management policy. The guideline is intended for assessment technical support units. 

### Template Email

*This formal email could be sent after initial more informal contact is made, as appropriate.*

Dear `r colorize("[contributing author name]", "red")`,

It is my pleasure to invite you to be a Contributing Author for the assessment of `r colorize("[assessment name]", "red")`, that is being developed by the [Intergovernmental Platform on Biodiversity and Ecosystem Services (IPBES)](https://ipbes.net/). Specifically, you are invited to contribute to Chapter `r colorize("[#]", "red")` on `r colorize ("[Chapter title]", "red")`. Your contribution would focus on the section on `r colorize("[topic]", "red")`, and you would be working with `r colorize("[IPBES authors names]", "red")`, who are `r colorize("[Lead Author(s)/Coordinating Lead Author(s)]", "red")` of the assessment and copied on this email.
  
The assessment of `r colorize("[assessment name]", "red")` was approved by the IPBES plenary in `r colorize("[year]", "red")` and will run until `r colorize("[year]", "red")`. It is  being developed based on the `r colorize("attached", "red")` scoping document.
  
As a contributing author, you would be linked directly to authors of the assessment and would be requested to provide specific contributions around themes, topics or geographic areas, as discussed with `r colorize ("[author name]", "red")`. Contributing Authors do not however have access to drafts of the full chapter due to IPBES rules on confidentiality of assessment drafts.
  
Contributions by contributing authors are always highly valuable and gratefully received, and support the development of key themes in an assessment chapter. However, we cannot guarantee that contributions will be reproduced in their entirety in the assessment. They may be edited during rounds of drafting and external review processes so that they form a part of a coherent chapter. Also, depending on space and chapter structures, the chapter may eventually use a synthesised version of your contribution. However, your full contribution may be made publicly available in a data management report or in supplementary materials that are available online as companion resources to the assessment.
  
If you choose to share any previously unpublished indigenous and local knowledge or materials, you will be asked to  confirm that you are authorised to share this information and confirm that it is non-confidential as specified in our guidance on documenting indigenous and local knowledge (https://ict.ipbes.net/ipbes-ict-guide/data-management/technical-guidelines/ILK-considerations).
  
To enable open science and accessibility within IPBES, the Platform has a data and knowledge management policy available at https://doi.org/10.5281/zenodo.3551078. 
    
Contributing authors are acknowledged on the first page of the Chapter to which they have contributed, but they are not listed as authors or within the citation of the chapter and assessment.
  
Finally, further communications guiding your contributions will be made directly by `r colorize("[author name]", "red")`. However, if you have any further questions feel free to contact me at `r colorize("[email]", "red")`.
  
We thank you for considering this invitation to contribute to this assessment, and we look forward to your valuable input.

`r colorize("[Signature]", "red")`